Three-Dimensional Temperature Characterization of Electrokinetically Pumped Microchannels

### 2018-11-29 paper.tex
agrego resumen conclusines y modificaciones

\title{Confocal Raman Thermometer for Microfluidic Devices}
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.

@@

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
              \and
              Juan Martin Cabaleiro \at
           Laboratorio de Fluidodinámica, FIUBA, CONICET, Av. Paseo Colón 850, Buenos Aires, Argentina, and Laboratorio de Micro y Nanofluídica y Plasmas, UdeMM, Av. Rivadavia 2258, Buenos Aires, Argentina  \\
              Tel.: +54 11 5285 0467\\              
              \email{jmcabaleiro@gmail.com}  
}

@@

\begin{abstract}
\keywords{Raman thermometer \and electroosmotic flow \and spatial resolution}
\end{abstract}

@@

Typical characteristic dimensions in microfluidic devices are in the range $\sim10-100\, \mathrm{\mu m}$. In these devices it is generally necessary to either move the liquid or move species in solution from one point to another in the microfluidic network. To do so, one needs a pumping or species transport strategy. In many cases this is done via electroosmotic pumping or electrophoresis \cite{hunter2001,lyklema1995}. In any of these cases, an electric current is passed through the liquid, the applied power has to be dissipated. Depending on the device dissipation efficiency, the liquid temperature will rise more or less. In any case, as these devices are in often used in biological applications \cite{tian2008}, temperature is a critical parameter that has to be monitored and controlled. Moreover, even in non biological applications, the electroosmotic and electrophoretic mobility are a function of temperature \cite{tang2006} so that temperature rise may lead to unwanted pumping or electrophoretic behaviour. The readers are referred to the reviews by Xuan~\cite{xuan2008} and Cetin and Li~\cite{cetin2008} for a deeper description of joule heating in electrokinetic flows. 
Different techniques were developed to measure temperature inside microchannels. A common approach is to introduce in the flow a dye with a temperature dependent property~\cite{tang2006,ross2001,erickson2003,dye1,dye2}. The measurement of this property with the proper calibration is used as a thermometer. This is an invasive method, with the disadvantage that the dye can interfere with the chemical reactions occurring in the channel or vice versa. This makes the method not useful for online monitoring. Moreover, the measurement ``integrates'' the temperature in the channel height, not allowing for the measurement of temperature profiles normal to the channel bottom.

Walrafen et al \cite{walrafen1} discovered that the infrared spectrum of water shows temperature dependent qualities and measured the existence of an isosbestic point around 3425~$\mathrm{cm^{-1}}$. This molecular property can be sensed using Raman Scattering to get a temperature dependent signal that is intrinsic of water. This method was proposed and proved useful in microchannels \cite{raman1}\cite{raman2}\cite{raman3} where an epi-illumination microscope was adapted to measure Raman Scattering by exciting the water with a visible laser. As water absorption is low in the visible spectrum \cite{absorption} this method proved to be non-invasive. In this work we propose an improvement over these methods were spatial resolution is enhanced by using a confocal collection scheme. This is achieved by focusing the collected Raman signal in a optical fibre (instead of the traditional pinhole used in most confocal microscopes). Also, the use of a 405~nm excitation beam maximizes temperature signal as Raman emission increases strongly with frequency \cite{faris}, besides being a cheap wavelength. Finally, a dark field illumination scheme is used to get scattering or fluorescence images of the sample at the same time temperature measurements take place. This allowed to perform temperature maps that characterize the heat flow in the channel and the geometric and construction parameters that dominate heat dissipation. Our results show good agreement with those found in the literature.

@@

The final spatial scan is shown in figure \ref{fig:long} where a longitudinal scan was performed. Again, temperature shows a constant behavior along the channel, similar to what was shown in figure \ref{fig:temporal}. This is consistent with theoretical and experimental results \cite{jouleteorico,xuan2008} where temperature rise occurs in the very beginning of the channel and then a constant value is achieved. A simple calculation considering that all heat flows trough the floor of the channel (glass) can be useful to estimate the required temperature gradient needed to reach a stationary regime where the extracted heat compensates Joule effect. For a typical dissipated power of 40 mW and a channel of dimensions $b = h = 100$ $\mathrm{\mu m}$ and $L$ = 20~mm the required temperature difference between the ceiling and the floor is

\begin{equation}
\Delta T = \frac{Ph}{kLb} = 1.7\,\, \mathrm{K},
\end{equation}

where $P$ is Joule heat dissipated on the channel and $k$ is water thermal conductivity. This upper bound is already on the limits of what is measurable with the technique and then our measurements are consistent with the glass base of the channel dominating as dissipation mechanism.

@@

To confirm this hypothesis curves of temperature rise as function of dissipated power per unit glass area were measured for different channel geometries. The result of this experiment is plotted in figure \ref{fig:rectas}. A linear fit was performed for each data set. Results show that all different geometries have a similar slope, proving that glass area is the most relevant parameter defining heat dissipation out of our channels. This is also consistent with results from other groups \cite{dye1} where rectangular channels made completely out of PDMS exhibit higher temperatures and longitudinal variations with similar conductivity and applied field.

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{figs/rectas.eps}
\caption{Temperature rise as a function of dissipated power per unit glass area on four different channel geometries. A linear fit is performed in each data set in order to compare the slope corresponding to each channel. The similarity between the lines seems to indicate that the glass floor of the channel dominates as dissipation mechanism.\label{fig:rectas}}
\end{figure}

>>>>

\title{Three-Dimensional Temperature Characterization of Electrokinetically Pumped Microchannels}
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.

@@

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{guillermobrinatti@gmail.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
              \and
              Juan Martin Cabaleiro \at
           Laboratorio de Fluidodinámica, FIUBA, CONICET, Av. Paseo Colón 850, Buenos Aires, Argentina, and Laboratorio de Micro y Nanofluídica y Plasmas, UdeMM, Av. Rivadavia 2258, Buenos Aires, Argentina  \\
              Tel.: +54 11 5285 0467\\              
              \email{jmcabaleiro@gmail.com}  
}

@@

\begin{abstract}
A novel method for non invasive, three dimensional temperature characterization in microfluidic devices is presented. A specially designed confocal microscope was constructed and used to measure water temperature by sensing the Raman spectrum variations of the liquid. A fiber coupled laser beam is used to pump the sample with 25~mW of optical power at 405~nm. This together with the confocal character of the device allows a 0.8~K temperature resolution and a 9~$\mu$m axial resolution using a 1 second integration time and a high numeric aperture microscope objective. This features make temperature profiling in all dimensions possible, in contraposition with previous methods where the information present in the height of the channel is lost. Using this technique different geometries of PDMS microchannels sealed with a 150~$\mu$m glass coverslip where studied, showing that heat flow trough the glass is the dominating dissipation mechanism and defines the maximum temperature in the channel. The results show good agreement with previous work found in literature.
\keywords{Raman thermometer \and electroosmotic flow \and spatial resolution}
\end{abstract}

@@

Typical characteristic dimensions in microfluidic devices are in the range $\sim10-100\, \mathrm{\mu m}$. In these devices it is generally necessary to either move the liquid or move species in solution from one point to another in the microfluidic network. To do so, one needs a pumping or species transport strategy. In many cases this is done via electroosmotic pumping or electrophoresis \cite{hunter2001,lyklema1995}. In any of these cases, an electric current is passed through the liquid, the applied power has to be dissipated. Depending on the device dissipation efficiency, the liquid temperature will rise more or less. In any case, as these devices are often used in biological applications \cite{tian2008}, temperature is a critical parameter that has to be monitored and controlled. Moreover, even in non biological applications, the electroosmotic and electrophoretic mobility are a function of temperature \cite{tang2006} so that temperature rise may lead to unwanted pumping or electrophoretic behaviour. The readers are referred to the reviews by Xuan~\cite{xuan2008} and Cetin and Li~\cite{cetin2008} for a deeper description of joule heating in electrokinetic flows. 
Different techniques were developed to measure temperature inside microchannels. A common approach is to introduce in the flow a dye with a temperature dependent property~\cite{tang2006,ross2001,erickson2003,dye1,dye2}. The measurement of this property with the proper calibration is used as a thermometer. This is an invasive method, with the disadvantage that the dye can interfere with the chemical reactions occurring in the channel or vice versa. This makes the method not useful for online monitoring. Moreover, the measurement ``integrates'' the temperature in the channel height, not allowing for the measurement of temperature profiles normal to the channel bottom.

Walrafen et al \cite{walrafen1} discovered that the infrared spectrum of water shows temperature dependent qualities and measured the existence of an isosbestic point around 3425~$\mathrm{cm^{-1}}$. This molecular property can be sensed using Raman Scattering to get a temperature dependent signal that is intrinsic of water. This method was proposed and proved useful in microchannels \cite{raman1}\cite{raman2}\cite{raman3} where an epi-illumination microscope was adapted to measure Raman Scattering by exciting the water with a visible laser. As water absorption is low in the visible spectrum \cite{absorption} this method proved to be non-invasive. In this work we propose an improvement over these methods were spatial resolution is enhanced by using a confocal collection scheme. This is achieved by focusing the collected Raman signal in a optical fibre (instead of the traditional pinhole used in most confocal microscopes). This result in a device capable of three dimensional maping of the channel, allowing the study of thermal gradients in every direction. Also, the use of a 405~nm excitation beam maximizes temperature signal as Raman emission increases strongly with frequency \cite{faris}, besides being a cheap wavelength. Finally, a dark field illumination scheme is used to get scattering or fluorescence images of the sample at the same time temperature measurements take place. This allowed to perform temperature maps that characterize the heat flow in the channel and the geometric and construction parameters that dominate heat dissipation. Our results show good agreement with those found in the literature.

@@

The final spatial scan is shown in figure \ref{fig:long} where a longitudinal scan was performed. Again, temperature shows a constant behavior along the channel, similar to what was shown in figure \ref{fig:temporal}. This is consistent with theoretical and experimental results \cite{jouleteorico,xuan2008} where temperature rise occurs in the very beginning of the channel and then a constant value is achieved.

Acá la cuentita que habia hecho yo quedó obsoleta dado el modelo de Martín. 
%A simple calculation considering that all heat flows trough the floor of the channel (glass) can be useful to estimate the required temperature gradient needed to reach a stationary regime where the extracted heat compensates Joule effect. For a typical dissipated power of 40 mW and a channel of dimensions $b = h = 100$ $\mathrm{\mu m}$ and $L$ = 20~mm the required temperature difference between the ceiling and the floor is 
%
%\begin{equation}
%\Delta T = \frac{Ph}{kLb} = 1.7\,\, \mathrm{K},
%\end{equation}

%where $P$ is Joule heat dissipated on the channel and $k$ is water thermal conductivity. This upper bound is already on the limits of what is measurable with the technique and then our measurements are consistent with the glass base of the channel dominating as dissipation mechanism.

@@

To confirm this hypothesis curves of temperature rise as function of dissipated power per unit glass area were measured for different channel geometries. The result of this experiment is plotted in figure \ref{fig:rectas}. A linear fit was performed for each data set. Results show that all different geometries have a similar slope, proving that glass area is the most relevant parameter defining heat dissipation out of our channels. This is also consistent with theoretical and experimental results from other groups \cite{} where temperature rise between all PDMS and PDMS/glass channels was compared, obtaining a much lower heating effect in the latter.

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{figs/rectas.eps}
\caption{Temperature rise as a function of dissipated power per unit glass area on four different channel geometries. A linear fit is performed in each data set in order to compare the slope corresponding to each channel. The similarity between the lines seems to indicate that the glass floor of the channel dominates as dissipation mechanism. Ref er to table 1 for channel dimensions.\label{fig:rectas}}
\end{figure}

\section{Conclusions}

A novel technique was presented that proven useful to characterize Joule heating effects in eletrokinetically pumped microchannels with three-dimensional resolution. The thermometer, that works using the temperature dependence in the Raman spectrum of liquid water, improves from previous methods by using a confocal collection scheme. This allows a much higher spatial resolution, allowing thermal characterization in the direction normal to the channels base, which is impossible in other methods where the complete channel height is integrated.

 A temperature resolution of about 0.8 K was achieved in a one second integration time using 25~mW of optical power at the sample provided from a 405~nm fiber coupled laser beam. Temperature profiles were measured in different channel geometries with results compatible with a simple heat flow model. From this experiments we conclude that heat dissipation is occurring through the glass floor of the channel in accordance with theoretical and experimental results found in literature. As temperature rise is an important parameter to control in many of the applications of these devices, the method provides a complete three dimensional characterization of Joule heating effects that allows for improved channel design and temperature monitoring.
